Evaluating the Performance and Energy Efficiency of the COSMO-ART model system

### 2014-04-24 trunk/paper_enahpc14/Introduction.tex
adds a few changes

The extended atmospheric model  COSMO-ART is computationally much more
demanding than  COSMO since a  large number of additional  tracers and
processes  have  to  be  considered.   Thus  the  model  is  currently
severely-limited in  terms of applicability and expensive  in terms of
energy  consumption.   Although  COSMO  has recently  been  ported  to
GPUs  \citep{Gysi-2014, Lapillonne-2014} within  the framework  of the
High Performance and High  Productivity Computing (HP2C) Initiative to
optimise    it    for     computational    and    energy    efficiency
(\url{http://www.hp2c.ch/}), significant investments  in ART are still
required to take  it to a similar level.   The efficiency of COSMO-ART
is     being    addressed    in     the    EU     Exa2Green    project
(\url{http://exa2green.eu/}). The  ultimate goal of the  project is to
deliver a  prototype code, which  provides an energy efficiency  of at
least five times of the  baseline value.  Such an implementation would
allow  the  community  to  investigate critical  questions  at  higher
resolution  and   over  longer  periods,   at  reduced  cost   to  the
environment.

This  work is  organized as  follows: in  Sec.~\ref{sec:1} we  give an
overview  of  related  work,   then  in  Sec.~\ref{sec:2}  we  briefly
introduce  COSMO-ART and specify  its technical  setup related  to the
investigated  performance  and  energy  evaluation  methods.   In  the
subsequent Sec.~\ref{sec:3},  we describe different  power measurement
equipments as well as energy-aware techniques employed to conduct this
benchmarking  study.  Sec.~\ref{sec:4}  presents  the performance  and
energy  requirements  of the  baseline  on  various architectures  and
highlights  areas  where  improvements   will  be  necessary  for  the
subsequent baseline refactoring.   Finally, we conclude with practical
guidelines  to identifying  energy  saving opportunities  and give  an
outlook to future research in Sec.~\ref{concl}.

>>>>

The extended atmospheric model  COSMO-ART is computationally much more
demanding than  COSMO since a  large number of additional  tracers and
processes  have  to  be  considered.   Thus  the  model  is  currently
severely-limited in  terms of applicability and expensive  in terms of
energy  consumption.   Although  COSMO  has recently  been  ported  to
GPUs  \citep{Gysi-2014, Lapillonne-2014} within  the framework  of the
High    Performance   and    High   Productivity    Computing   (HP2C)
Initiative \citep{HP2C},  to optimise it for  computational and energy
efficiency, significant investments in  ART are still required to take
it to a similar level.  The efficiency of COSMO-ART is being addressed
in the  EU Exa2Green project \citep{EXA2GREEN} to  deliver a prototype
code, which  provides an energy efficiency  of at least  five times of
the baseline value.  Such  an implementation would allow the community
to investigate critical questions at higher resolution and over longer
periods, at reduced cost to the environment.

This  work is  organized as  follows: in  Sec.~\ref{sec:1} we  give an
overview  of  related  work,   then  in  Sec.~\ref{sec:2}  we  briefly
introduce  COSMO-ART and specify  its technical  setup related  to the
investigated   performance   and   energy  evaluation   methods.    In
Sec.~\ref{sec:3}   we  describe   HPC  platforms,   power  measurement
equipments   and  software  environment   employed  to   conduct  this
benchmarking study.  Sec.~\ref{sec:4}  presents performance and energy
requirements  of the  baseline on  these architectures  and highlights
areas where improvements will be necessary for the subsequent baseline
refactoring.   Finally,  we  conclude  with  practical  guidelines  to
identifying energy saving opportunities  and give an outlook to future
research in Sec.~\ref{concl}.

### 2014-04-24 trunk/paper_enahpc14/Sec3:Environment-setup.tex
adds a few changes

\subsubsection{CSCS - Monch}
Monch is  a 10 rack  NEC-provided and dual-socket Intel  Ivy Bridge-EP
based  cluster, utilised  by  scientists  who are  part  of the  Swiss
Platform     for      Advanced     Scientific     Computing     (PASC,
\url{http://www.pasc-ch.org/}). It is composed of 312 standard compute
nodes, 24 large-memory compute nodes and 24 huge-memory compute nodes.
Each standard  compute node comprises  two Intel Ivy  Bridge Efficient
Performance (EP)  E5-2660 v2 ten-core processors operating  at 2.2 GHz
base  clock speed,  themselves connected  by a  high  speed InfiniBand
network based on Mellanox SX6036  managed FDR switches, with a 56 Gb/s
speed.  Each core  has 32 KB instruction and 32 KB  data L1 caches and
256  KB of L2  cache. All  10 cores  share a  25 MB  L3 cache  and the
platform has  32GB of  DDR3 1600 MHz  RAM. For  our energy-to-solution
benchmark, a  full rack  of Monch constituted  of 52  standard compute
nodes (monchc[029-080]) was considered.

@@

\subsubsection{CSCS - E3METER metering products}
Supercomputer  clusters considered  for our  experiments at  the Swiss
National Supercomputing Center (CSCS)  of ETH Zurich are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks
(\url{http://riedonetworks.com/}),  that  enable  to monitor  and  log
power  consumption of  the  IT infrastructure  as  well as  constantly
analyze  line  voltage,  current,  power-factor,  frequency  with  1\%
accuracy.   Using reliable  narrowband  powerline communication  (PLC)
technology,  all metering  and power  quality data  from each  IPS are
centrally collected by the E3METER Data Concentrator, via the existing
power cables thus  avoiding the need for extra  cabling.  This data is
made  available  via SNMP,  HTTP,  TELNET  through  the built-in  Fast
Ethernet  port.   Time  synchronisation  is guaranteed  by  using  NTP
servers.   Measured data  is accessed  through the  open  source Cacti
software including  the E3METER  Cacti Plugin to  scan the  entire PLC
network and monitor  in real-time the power usage  of individual rack,
recorded in 5 minutes interval periods.

>>>>

\subsubsection{CSCS - Monch}
Monch is  a 10 rack  NEC-provided and dual-socket Intel  Ivy Bridge-EP
based  cluster, utilised  by  scientists  who are  part  of the  Swiss
Platform for Advanced Scientific  Computing (PASC) \citep{PASC}. It is
composed of 312 standard  compute nodes, 24 large-memory compute nodes
and  24  huge-memory  compute   nodes.   Each  standard  compute  node
comprises two  Intel Ivy Bridge Efficient Performance  (EP) E5-2660 v2
ten-core processors operating at  2.2 GHz base clock speed, themselves
connected by a high speed  InfiniBand network based on Mellanox SX6036
managed  FDR switches,  with a  56 Gb/s  speed.  Each  core has  32 KB
instruction and 32  KB data L1 caches  and 256 KB of L2  cache. All 10
cores share a  25 MB L3 cache  and the platform has 32GB  of DDR3 1600
MHz RAM.  For our energy-to-solution  benchmark, a full rack  of Monch
constituted  of  52   standard  compute  nodes  (monchc[029-080])  was
considered.

@@

\subsubsection{CSCS - E3METER metering products}
Supercomputer  clusters considered  for our  experiments at  the Swiss
National Supercomputing Center (CSCS)  of ETH Zurich are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks
\citep{Riedonetworks},   that  enable   to  monitor   and   log  power
consumption  of the IT  infrastructure as  well as  constantly analyze
line  voltage,  current, power-factor,  frequency  with 1\%  accuracy.
Using  reliable narrowband  powerline communication  (PLC) technology,
all  metering and  power  quality  data from  each  IPS are  centrally
collected  by the E3METER  Data Concentrator,  via the  existing power
cables thus  avoiding the need for  extra cabling.  This  data is made
available via  SNMP, HTTP, TELNET  through the built-in  Fast Ethernet
port.   Time  synchronisation  is  guaranteed by  using  NTP  servers.
Measured  data is  accessed  through the  open  source Cacti  software
including the E3METER Cacti Plugin  to scan the entire PLC network and
monitor in real-time the power usage of individual rack, recorded in 5
minutes interval periods.
